\documentclass{ctuthesis}

\input{src/preamble.tex}

\input{src/abstract.tex}

\begin{document}

\maketitle

\begin{enumerate}
	\item Background (SoTA) overview
		\begin{itemize}
			\item Graph learning intro
				\begin{itemize}
					\item Why graphs?
					\item Tasks
						\begin{itemize}
							\item Node classification
							\item Link prediction
							\item Graph classification
						\end{itemize}
					\item Transductive vs. inductive tasks
					\item \cite{gori_new_2005} - overview
				\end{itemize}
			\item Random walk based models
				\begin{itemize}
					\item DeepWalk
					\item LINE?
					\item Node2Vec
				\end{itemize}
			\item MPNNs / convolutional models
				\begin{itemize}
					\item Introduction to message passing
					\item GCN
					\item GraphSAGE
				\end{itemize}
			\item WL test and expressivity in GNNs
				\begin{itemize}
					\item The WL test
					\item The WL test and graph convolution
					\item GIN?
					\item Oversmoothing in graphs
					\item Oversquashing in graphs
					\item The 4 Gs?
				\end{itemize}
			\item Explainability on graphs
				\begin{itemize}
					\item Overview of graph explainability
					\item Instance-level post-hoc explainability
						\begin{itemize}
							\item Gradient-based methods
								\begin{itemize}
									\item Grad-CAM
								\end{itemize}
							\item Surrogate methods
								\begin{itemize}
									\item GraphLime
								\end{itemize}
							\item Perturbation-based methods
								\begin{itemize}
									\item GNNExplainer
									\item PGExplainer
									\item SubgraphX
								\end{itemize}
						\end{itemize}
					\item Model-level explainability
						\begin{itemize}
							\item XGNN
							\item GraphChef?
						\end{itemize}
				\end{itemize}
		\end{itemize}
	\item My research so far
		\begin{itemize}
			\item Graph coarsening for efficient GNNs -- HARP based approach (\cite{dedic_balancing_2023, dedic_balancing_2024})
			\item Graph coarsening for efficient GNNs -- direct edge collapsing (\cite{prochazka_scalable_2022})
			\item Graph properties and their role in classification and its performance (\cite{prochazka_which_2023})
			\item Maybe something more recent with explainability?
		\end{itemize}
	\item Future ideas and intended directions
		\begin{itemize}
			\item Graph explainability
			\item GNN HPO
			\item Simple graph models such as RMG?
		\end{itemize}
\end{enumerate}

\printbibliography

\end{document}
